---
title: 'Statistical report: Prognostic impact of non-cardiac comorbidity burden in patients hospitalized for AHF and stratified by EF categories: rEF, mrEF and pEF'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  pdf_document:
    dev: cairo_pdf
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    number_sections: yes
link-citations: yes
bibliography: references.bib
nocite: '@*'
urlcolor: blue
linkcolor: black
---

\newpage 
\tableofcontents 
\listoftables
\listoffigures
\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

cacheon <- TRUE
```             

# Data handling

## Data source

hf3_lt_fu_data_soladis_jan19.sas7bdat (from the folder DATABASE_ANALYZED.ZIP). 

## Inclusion/exclusion criteria

```{r flow}
default_kable(flow, caption = "Flowchart")
```

First patient in: `r min(edata$num_dmVisitdt)` and last patient in: `r max(edata$num_dmVisitdt)`. 

The median age (IQR) is `r edata %>% summarise(med = fn(median(num_age, na.rm = T), dig = 1),
                                             q1 = fn(quantile(num_age, na.rm = T, probs = 0.25), dig = 1),
                                             q3 = fn(quantile(num_age, na.rm = T, probs = 0.75), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
                                   pull(out)` and 
`r edata %>% count(num_dmgender) %>%
  mutate(perc = fn(n / sum(n) * 100, 1)) %>%
  filter(num_dmgender == "Female") %>%
  pull(perc)`% females.  

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with some care.

EF is defined as EF by Echo-Doppler recorded at discharge, or if this is missing, Last known EF recorded in Characteristics. 

The number of non-cardiac comorbidities score consist of the following: `r paste(comorbs, collapse = ", ")`.

## Missing data

For patients with missing information on the date of hospitalisation, 
the time to hospitalisation was imputed with half the time to last
follow-up. Further, times to hospitalisation that were larger than time 
to death or last follow-up were set to death or last follow-up. 

## Baseline characteristics

```{r, child = "../src/tab1.Rmd"}

```

\newpage

## Comorbs/NYHA

```{r, child = "../src/nyhaplot.Rmd"}

```

\newpage

## Outcome analysis

The following outcomes are considered: 

- NYHA at admission (all population)
- NYHA at discharge (discharge population)
- Change in NYHA (discharge population)
- In-hospital mortality (all population)
- Length of stay at index hospitalization >= 7 days (all population)
- 1 year (approx) All-cause mortality (long-term outcome population)
- 1 year (approx) All-cause hospitalisations (long-term outcome population)
- 1 year (approx) CV mortality (long-term outcome population)
- 1 year (approx) HF hospitalisations (long-term outcome population)
- 1 year (approx) All-cause mortality/HF hospitalisations (long-term outcome population)


For all models adjustment is performed for `r paste(modvars, collapse = ", ")` where 
the continuous variables are modelled using cubic splines with 4 degrees of freedom. 

NYHA class at admission and discharge and length of stay were categorized as 
dichotomous variables. 

For the binomial outcomes (NYHA, LOS and in-hospital mortality) 
a generalized linear mixed-effects model (GLMM) with a logit link and a random effect for country was used. 

The long-term outcomes are presented with cumulative incidence curves. 
Cox proportional hazards regressions with a frailty term for country were 
used to model the time to first event analysis. 

As a consistency analysis the long-term outcomes (not all-cause death)
was modelled using a sub-distributional hazards model [@fg] where death was treated as a competing event. 

The long-term outcomes were also modelled, as above, 
separately in each EF group and presented in Forest plots. Further, they were 
analysed with each non-cardiac 
comorbity individually for the whole population and separately in each EF group and presented in Forest plots. 

Time is from date of discharge, or if this is missing, date of admission and 
censored at death not defined as an event or follow-up.  

The median (min-max) follow-up for the long-term outcomes is 
`r edata %>% filter(survpop) %>% summarise(med = fn(median(outtime_death / 365.25 * 12), dig = 1),
                                             min = fn(min(outtime_death / 365.25 * 12), dig = 1),
                                             max = fn(max(outtime_death / 365.25 * 12), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", min, "-", max, ")")) %>%
                                   pull(out)` months for a total of 
                                   `r edata %>% filter(survpop) %>% 
                                   summarise(sumpy = fn(sum(outtime_death) / 365.25, dig = 0)) %>%
                                   pull(sumpy)` patient-years of follow-up.
                                   
### Assumptions

The proportional hazards assumption was investigated using the scaled Schoenfeld 
residuals (cox.zph in [@survival-package]) for the survival outcomes. 

\newpage
                                   
```{r, child = "../src/outtab.Rmd"}

```

\newpage

```{r, child = "../src/noforest.Rmd"}

```

\newpage

```{r, child = "../src/indforest.Rmd"}

```

\newpage

```{r, child = "../src/km.Rmd"}

```

\newpage

# Reproducibility

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: 
https://github.com/KIHeartFailure/esccomorb. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References
