---
title: 'Statistical report: Prognostic impact of non-cardiac comorbidity burden in patients hospitalized for AHF and stratified by EF categories: rEF, mrEF and pEF'
subtitle: 'DRAFT'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    number_sections: yes
link-citations: yes
bibliography: references.bib
nocite: '@*'
urlcolor: blue
linkcolor: black
header-includes:
   - \usepackage{draftwatermark}
---

\newpage 
\tableofcontents 
\listoftables
\listoffigures
\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

cacheon <- TRUE
```             

# Data handling

## Data source

hf3_lt_fu_data_soladis_jan19.sas7bdat (from the folder DATABASE_ANALYZED.ZIP). 

## Inclusion/exclusion criteria

```{r flow}
default_kable(flow, caption = "Flowchart")
```

First patient in: `r min(edata$num_dmVisitdt)` and last patient in: `r max(edata$num_dmVisitdt)`. 

The median age (IQR) is `r edata %>% summarise(med = fn(median(num_age, na.rm = T), dig = 1),
                                             q1 = fn(quantile(num_age, na.rm = T, probs = 0.25), dig = 1),
                                             q3 = fn(quantile(num_age, na.rm = T, probs = 0.75), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
                                   pull(out)` and 
`r edata %>% count(num_dmgender) %>%
  mutate(perc = fn(n / sum(n) * 100, 1)) %>%
  filter(num_dmgender == "Female") %>%
  pull(perc)`% females.  

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with some care.

EF is defined as EF by Echo-Doppler recorded at discharge, or if this is missing, Last known EF recorded in Characteristics. 

Number of non-cardiac comorbidities consist of the following: `r paste(comorbs, collapse = ", ")`

## Missing data

For patients with missing information on the date of hospitalisation, 
the time to hospitalisation was imputed with half the time to last
follow-up. Further, times to hospitalisation that were larger than time 
to death or last follow-up were set to death or last follow-up. 

Might do this: Missing data for the covariates included in the models was imputed with multiple 
imputation using mice [@mice] for 10 datasets and 10 iterations. 
Variables included in the imputation model are the same as are 
included in the cox regression. Death was included as the Nelson-Aalen estimator 
in the imputation model. However, will depend on what variables are adjusted for in the models. 
Currently Age, sex, SBP at admission, ischemic aetiology, atrial fibrillation, history of MI and there is 
 very little missing for these so not really justifiable to do mi. 

## Baseline characteristics

```{r, child = "../src/tab1.Rmd"}

```

\clearpage
\newpage

## Comorbs/NYHA

```{r, child = "../src/nyhaplot.Rmd"}

```

\clearpage
\newpage

## Outcome analysis

The following outcomes are considered: 

- NYHA at admission
- NYHA at discharge
- Change in NYHA
- In-hospital mortality
- Length of stay
- 1 year (approx) All-cause mortality
- 1 year (approx) All-cause hospitalisations
- 1 year (approx) CV mortality
- 1 year (approx) HF hospitalisations
- 1 year (approx) All-cause mortality/HF hospitalisations

For all models adjustment is performed for `r paste(modvars, collapse = ", ")`. 

For the binomial outcomes (decrease in NYHA and in-hospital mortality) 
a generalized linear mixed-effects model (GMMM) with a logit link and a random effect for country was used. 

For the ordinal outcomes (NYHA class at admission and discharge) 
a Cumulative Link Mixed Model with a random effect for country was used. 

Length of stay is modelled using a generalized linear mixed-effects model 
(GLMM) for the negative binomial family with a random effect for country. 
(I am assuming LOS is lenght of stay at the index hosp, not during follow-up???)

The long-term outcomes are presented with the Kaplan-Meier curves 
(for outcomes with all-cause death) or cumulative incidence curves (with death as a competing risk). 
Cox proportional hazards regressions with a frailty term for country were 
used to model the time to first event analysis. 

As a consistency analysis the long-term outcomes (not all-cause death)
was modelled using a sub-distributional hazards model [@fg] where death was treated as a competing event. 

The long-term outcomes were also modelled in a similar manner as above but with 
an interaction term for EF, thereby enabling separate estimates for each EF group, and presented in a Forest plot. 
The models were adjusted for the same variables as above. 

Time is from date of discharge, or if this is missing, date of admission and 
censored at death not defined as an event or follow-up.  

The median (min-max) follow-up for the long-term outcomes is 
`r edata %>% filter(survpop) %>% summarise(med = fn(median(outtime_death / 365.25 * 12), dig = 1),
                                             min = fn(min(outtime_death / 365.25 * 12), dig = 1),
                                             max = fn(max(outtime_death / 365.25 * 12), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", min, "-", max, ")")) %>%
                                   pull(out)` months for a total of 
                                   `r edata %>% filter(survpop) %>% 
                                   summarise(sumpy = fn(sum(outtime_death) / 365.25, dig = 0)) %>%
                                   pull(sumpy)` patient-years of follow-up.
                                   
### Assumptions

The proportional hazards assumption was investigated using the scaled Schoenfeld 
residuals (cox.zph in [@survival-package]) for the primary outcome. 
Possible outliers were visually inspected by plotting the dfbetas. 
=> XXXX. 

\clearpage
\newpage
                                   
```{r, child = "../src/outtab.Rmd"}

```

\clearpage
\newpage

```{r, child = "../src/outforestef.Rmd"}

```

\clearpage
\newpage

```{r, child = "../src/km.Rmd"}

```

\clearpage
\newpage

# Reproducibility

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: 
https://github.com/KIHeartFailure/esccomorb. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References